\textcolor{red}{Disambiguating Monocular Depth Estimation with a Single Transient}

### 2020-02-28 paper/eccv2020/main.tex
Updated with eccv format, initial revisions.
Moving figures around, adding edit red text.
pass over intro

% INITIAL SUBMISSION - The following two lines are NOT commented
% CAMERA READY - Comment OUT the following two lines
% \usepackage{ruler}
% \usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}

@@

\def\eg{\emph{e.g.}} \def\Eg{\emph{E.g.}}
\def\ie{\emph{i.e.}} \def\Ie{\emph{I.e.}}
\def\cf{\emph{c.f.}} \def\Cf{\emph{C.f.}}
\def\etc{\emph{etc}} \def\vs{\emph{vs.}}
\def\wrt{w.r.t.} \def\dof{d.o.f.}
\def\etal{\emph{et al.}}
\graphicspath{{figures/}{"G:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"H:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"/Users/gordon/mount/teamdrive/Shared
    drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}
  {"/Volumes/GoogleDrive/Shared drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}}
% Pages are numbered in submission mode, and unnumbered in camera-ready

@@

\title{Improving Monocular Depth Estimation with Global Depth Histogram
  Matching using a Single SPAD Transient} % Replace with your title

@@

\begin{abstract}
  Existing monocular depth estimation algorithms successfully predict the
  relative depth {order} of objects in a scene.
  However, because of the fundamental scale ambiguity associated with monocular
  images, these algorithms fail at correctly predicting an object's true metric
  depth.
  In this work, we demonstrate how a depth histogram of the scene, which can be
  readily captured using a single-pixel diffused single-photon avalanche diode
  (SPAD), can be fused with the output of existing monocular depth estimation
  algorithms to resolve the depth ambiguity problem.
  We validate this novel sensor fusion technique experimentally and in extensive
  simulation. We show that it dramatically improves the performance of several
  state-of-the-art monocular depth estimation algorithms.
  \dots
  \keywords{monocular depth estimation, single-photon avalanche diode, histogram
  matching, LiDAR}
\end{abstract}

@@

\clearpage
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{references}
\end{document}

>>>>

% INITIAL SUBMISSION - The following two lines are NOT commented
% CAMERA READY - Comment OUT the following two lines
\usepackage{ruler}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}

@@

\def\eg{\emph{e.g.}} \def\Eg{\emph{E.g.}}
\def\ie{\emph{i.e.}} \def\Ie{\emph{I.e.}}
\def\cf{\emph{c.f.}} \def\Cf{\emph{C.f.}}
\def\etc{\emph{etc}} \def\vs{\emph{vs.}}
\def\wrt{w.r.t.} \def\dof{d.o.f.}
\def\etal{\emph{et al.}}

\newcommand{\edit}[1]{\textcolor{red}{#1}}

\graphicspath{{figures/}{"G:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"H:/Shared drives/Stanford Computational
    Imaging/Projects/single_spad_depth/figures/"}{"/Users/gordon/mount/teamdrive/Shared
    drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}
  {"/Volumes/GoogleDrive/Shared drives/Stanford Computational Imaging/Projects/single_spad_depth/figures/"}}
% Pages are numbered in submission mode, and unnumbered in camera-ready

@@

%\title{Improving Monocular Depth Estimation with Global Depth Histogram Matching using a Single SPAD Transient} % Replace with your title
\title{\textcolor{red}{Disambiguating Monocular Depth Estimation with a Single Transient}} % Replace with your title

@@

\begin{abstract}
  Monocular depth estimation algorithms successfully predict the
  relative depth {order} of objects in a scene.
  However, because of the fundamental scale ambiguity associated with monocular
  images, these algorithms fail at correctly predicting true metric
  depth. In this work, we demonstrate how a depth histogram of the scene, which can be readily captured using a \textcolor{red}{single-pixel time-resolved detector}, can be fused with the output of existing monocular depth estimation algorithms to resolve the depth ambiguity problem.
  We validate this novel sensor fusion technique experimentally and in extensive
  simulation. We show that it significantly improves the performance of several
  state-of-the-art monocular depth estimation algorithms.  
  \keywords{depth estimation, time-of-flight imaging}
\end{abstract}

@@

\clearpage
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{references}

\end{document}

### 2020-02-28 paper/eccv2020/sections/introduction.tex
pass over intro

Estimating dense 3D geometry from 2D images is an important open problem with
applications to robotics, autonomous driving, and medical imaging. Depth maps
are a common representation of scene geometry and are useful precursors to
higher-level scene understanding tasks such as pose estimation and object
detection. Additionally, many computer vision tasks rely on depth sensing,
including navigation~\cite{geiger2013vision}, semantic
segmentation~\cite{gupta2013perceptual,ren2012rgb,silberman2012indoor}, 3D
object
detection~\cite{gupta2014learning,lin2013holistic,shrivastava2013building,song2014sliding,song2016deep},
and 3D object
classification~\cite{maturana2015voxnet,qi2016volumetric,wu20153d}.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.6\columnwidth]{teaser.pdf}
  \caption{Monocular depth estimation predicts a depth map (upper right) from a
    single RGB image (upper left). The ill-posedness of the problem prevents
    reliable absolute depth estimation, resulting in large errors (inset image).
    The proposed method uses a measurement from a diffused SPAD (lower left),
    which resembles the shape of a histogram of the ground truth depth map, to
    correct the output of the depth estimation and optimize the quality of the
    estimated absolute depth (lower right).}
  \label{fig:teaser}
\end{figure}

@@

Although access to the median ground truth depth is impossible in a realistic
scenario, low-cost sensors capable of capturing aggregated depth information
from a scene are readily available. For example, the proximity sensor on recent
generation Apple iPhones uses a low-power pulsed light source and a
single-pixel, single-photon avalanche diode (SPAD) to sense distance to an
object directly in front of the phone. SPADs have also been used to recover full
3D geometry, forming the backbone of emerging single-photon LiDAR
systems~\cite{Kirmani:2014,Li:2019,pawlikowska2017single}. However, SPADs have
not been used for 3D imaging on consumer electronics because the requirement for
ultra-fast timing electronics makes it difficult to produce high-resolution
arrays at low cost, and because the scanning requirement for single-pixel
systems introduces a point of mechanical failure and complicates
high-resolution, high-framerate imaging.

@@

Here, we propose to use a single-pixel SPAD and pulsed light source in an
unconventional way: rather than optically focusing them to record the distance
to a single scene point, we diffuse the emitted light and the detector over the
entire scene. The resulting measurement resembles a histogram of the depth of
the scene and we demonstrate that this can be used to achieve accurate absolute
depth when combined with a monocular depth estimate
(\cf~Figure~\ref{fig:teaser}).

@@

To this end, we develop a sensor fusion strategy that processes the ordinal
depth computed by a monocular depth estimator to be consistent with the
measurements captured by the diffused SPAD. We demonstrate in extensive
simulations that our approach achieves substantial improvements in the quality
of the estimated depth maps, regardless of which specific depth estimator is
used. \edit{Moreover, we build an RGB-SPAD camera prototype that takes scanned
measurements of a scene and digitally sums them to approximate the effect of a
diffused laser and SPAD. While this approximation is not perfect, it is
sufficient to demonstrate that our sensor
fusion strategy can produce higher-quality
depth maps for a variety of captured scenes compared to MDE alone.}  With this
work, we present a practical way to disambiguate depth estimation with RGB
images using minimal additional sensing hardware. Specifically, we make the
following contributions:

%
\begin{itemize}
	\item We propose augmenting an RGB camera with a global depth histogram captured by a diffused SPAD to address scale ambiguity error in monocular depth estimators.	
  \item We analyze this approach on indoor scenes using the NYU Depth v2 dataset and demonstrate that our approach is able to resolve scale ambiguity while being fast and easy to implement.
	\item We build a prototype RGB-SPAD camera and evaluate its efficacy on captured data, assessing both the quality and the ability of our method to help generalization of monocular depth estimators across scene types. 
\end{itemize}

\edit{Our method is not without limitations. It still requires a laser and
single-pixel LiDAR detector and is sensitive to ambient photons, which limit the
maximum range.
Being a variant of histogram matching, our method is
not able to resolve ordinal depth errors (errors where an object is wrongly
placed closer or farther relative to another object). In spite of these
limitations, we find that our method performs remarkably well in practice.}

>>>>

Estimating dense 3D geometry from 2D images is an important problem with
applications to robotics, autonomous driving, and medical imaging. Depth maps
are a common representation of scene geometry and are useful precursors to
higher-level scene understanding tasks such as pose estimation and object
detection. Additionally, many computer vision tasks rely on depth sensing,
including navigation~\cite{geiger2013vision}, semantic
segmentation~\cite{gupta2013perceptual,ren2012rgb,silberman2012indoor}, 3D
object
detection~\cite{gupta2014learning,lin2013holistic,shrivastava2013building,song2014sliding,song2016deep},
and 3D object
classification~\cite{maturana2015voxnet,qi2016volumetric,wu20153d}.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.3\columnwidth]{teaser.pdf}
  \caption{Monocular depth estimation predicts a depth map (upper right) from a
    single RGB image (upper left). The ill-posedness of the problem prevents
    reliable absolute depth estimation, resulting in large errors (inset image).
    The proposed method uses a \edit{single transient measurement aggregating the time-of-flight information of the entire scene (lower left)} to correct the output of the depth estimation and optimize the quality of the
    estimated absolute depth (lower right).}
  \label{fig:teaser}
\end{figure}

@@

Although access to the median ground truth depth is impossible in a realistic
scenario, low-cost sensors capable of capturing aggregated depth information
from a scene are readily available. For example, the proximity sensor on recent
generation Apple iPhones uses a low-power pulsed light source and a
single-pixel \edit{time-resolved detector} %, single-photon avalanche diode (SPAD) 
to sense distance to an
object directly in front of the phone. %SPADs 
\edit{
Time-resolved detectors, such as avalanche photon diodes (APDs) or single-photon avalanche diodes (SPADs), measure the full waveform of time-resolved incident radiance at each pixel. These detectors form the backbone of modern LiDAR
systems~\cite{Kirmani:2014,Li:2019,pawlikowska2017single}. However, single-photon sensor arrays} have
not been used for 3D imaging on consumer electronics because the requirement for
ultra-fast timing electronics makes it difficult to produce high-resolution
arrays at low cost, and because the scanning requirement for single-pixel
systems introduces a point of mechanical failure and complicates
high-resolution, high-framerate imaging.

@@

Here, we propose to use a single-pixel \edit{time-resolved detector} and pulsed light source in an
unconventional way: rather than optically focusing them to record the distance
to a single scene point, we diffuse the emitted light and \edit{aggregate the reflected light over the
entire scene with the detector}. The resulting \edit{transient} measurement resembles a histogram of the depth of
the scene and we demonstrate that this can be used to achieve accurate absolute
depth when combined with the estimate of \edit{any monocular depth estimator in a post-processing step} (\cf~Figure~\ref{fig:teaser}).

@@

To this end, we develop a sensor fusion strategy that processes the ordinal
depth computed by a monocular depth estimator to be consistent with the
measurements captured by the \edit{aggregated time-resolved} detector. We demonstrate in extensive
simulations that our approach achieves substantial improvements in the quality
of the estimated depth maps, regardless of which specific depth estimator is
used. \edit{Moreover, we build an camera prototype that combines an RGB camera and a single-pixel time-resolved detector.}  With this
work, we present a practical way to disambiguate depth estimation with RGB
images using minimal additional sensing hardware. Specifically, we make the
following contributions:
%
\begin{itemize}
	\item We propose augmenting an RGB camera with a global depth histogram \edit{aggregated by a diffused time-resolved detector} to address scale ambiguity error in monocular depth estimators.	
  \item We analyze this approach on indoor scenes using the NYU Depth v2 dataset and demonstrate that our approach is able to resolve scale ambiguity while being fast and easy to implement.
	\item We build a prototype camera and evaluate its efficacy on captured data, assessing both the quality and the ability of our method to help generalization of monocular depth estimators across scene types. 
\end{itemize}

\edit{\emph{Overview of Limitations:} our prototype camera uses a scanned SPAD and digitally aggregates the captured transients to emulate a single diffused measurement. The benefit of this approach is that we have access to ground truth depth, allowing us to evaluate the efficacy of our method with measured data. However, when operating SPADs in certain conditions, they may observe a nonlinear aggregation effect known as pileup. In these conditions, the emulated measurements may differ from a single optically diffused measurement. Yet, we experimentally verify that digitally and optically aggregated measurements captured with our system are very similar and pileup could also be computationally corrected~\cite{Heide:2018,Rapp:2019}, although we did not attempt to do this.}

%\edit{Our method is not without limitations. It still requires a laser and single-pixel LiDAR detector and is sensitive to ambient photons, which limit the maximum range. Being a variant of histogram matching, our method is not able to resolve ordinal depth errors (errors where an object is wrongly placed closer or farther relative to another object). In spite of these limitations, we find that our method performs remarkably well in practice.}
